\chapter*{Acknowledgments}

Throughout this thesis, I have received a lot of support and assistance.

First of all, I would like to thank Pr. Yves Le Traon who allowed me to
pursue my PhD studies in his group and under his supervision. He believed in my research and provided me with valuable feedback. I am also grateful for the opportunities of teaching and the confidence he placed in me when organizing academic activities.

Of all the people in my thesis, I am especially grateful to my co-supervisors Mike Papadakis and Marinos Kintis for their patience, their advice, their training, as well as their support. They taught me how to conduct proper research and present my results to others. The frequent discussions we had allowed me to better understand the field, but also to become a better researcher altogether.

I am also thankful to my co-authors for their efforts and feedback that contributed to making this thesis stronger. I also want to thank all jury members for their interest in my research and the time invested in my dissertation.

I would like to express my gratitude to all my colleagues from SERVAL (SnT) for all the good discussions we had but also for their support during the darker times of my thesis.

I am thankful to the National Research Fund (FNR) of Luxembourg for trusting in the initial ideas that lead to this dissertation by granting the funding that supported my thesis. Equally, I would like to thank BGL BNP Paribas, and especially the test factory team, who was at my side for four years and offered me feedback and continuous trust.

Finally and more personally, I thank my partner Oksana Stolietnia, for her love, constant support, and patience during my thesis. I also thank my mother and my sister for their unconditional support and help throughout this adventure.